---
title: "p8105_hw3_tw2918"
output: github_document
date: "2023-10-07"
---

```{r, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## 1. Load Data

```{r}
library(p8105.datasets)
data("instacart")
```

## 2.  write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations.

* The dataset contains cleaned and limited data of the “The Instacart Online Grocery Shopping Dataset 2017”, which records theover 3 million online grocery orders from more than 200,000 Instacart users. 
*The dataset contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each `row`/observation in the dataset is a product from an order.
* Some variables reflect key identifications of the orders and customers including `order_id` and `order_id`. There is a single order per consumers in this dataset.
* `eval_set`reflects evaluation set this order belongs in, which is exclusively "`train`" in this dataset.
* `order_number` reflects the order sequence number for this user. Its value ranges from `r min(instacart$order_number)` to`r max(instacart$order_number)`. The average order_number is `r mean(instacart$order_number)`. 
* Some variables reflect key information of the products, including `product_id` and `product_name`, which aisle and department the prodcuts are purchased from, including  `aisle_id`,  `aisle`,  `department_id` and  `department`, and the order in which each product was added to cart `add_to_cart_order` , from `r min(instacart$add_to_cart_order)` to `r max(instacart$add_to_cart_order)`. 
* `reordered` variable is a binary indicator off this prodcut has been ordered by this user in the past, 0 otherwise. 
* The rest of the varialbes reflect date and time of the order, including the day of the week on which the order was placed `order_dow` ,  the hour of the day on which the order was placed, `order_hour_of_day`, and days since the last order, capped at `30`, `NA` if `order_number=1`. The average days since the last order is `r mean(instacart$order_hour_of_day)`. 

## 3. Answer questions and comments on result

### Calculate number of aisle in dataset and which aisles is the most items ordered from

```{r}
mode = function(x, na.rm = FALSE) {
  if(na.rm){ #if na.rm is TRUE, remove NA values from input x
    x = x[!is.na(x)]
  }
  val <- unique(x)
  return(val[which.max(tabulate(match(x, val)))])
}

mode(instacart$aisle_id)
mode(instacart$aisle)
```

* There are `r max(instacart$aisle_id)`  aisles in the dataset, and aisle `r mode(instacart$aisle_id)`, `r mode(instacart$aisle)` aisles is the most items ordered from. 


### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |>
  count(aisle)|>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n )) +
  geom_point() +
  labs(
    title = "number of items ordered in each aisle",
    x = "Aisle Name",
    y = "Number of items ordered",
  ) +
  scale_y_continuous(
    limits = c(10000, 160000)
  )+
  theme(axis.text.x =element_text(angle = 60, hjust = 1))
```


### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  select(aisle, product_name, everything()) |>
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name)|>
  mutate(rank = min_rank(desc(n)))|>
  filter(rank< 4)|>
  arrange(desc(n))|>
  knitr::kable()
```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  select (product_name, order_hour_of_day, order_dow)|>
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) |>
  mutate(
    dow = recode(order_dow, "0" = "day_0", "1" = "day_1", "2" = "day_2", "3" = "day_3", "4" = "day_4", "5" = "day_5", "6" = "day_6"))|>
  group_by(product_name, dow)|>
  summarize(mean_hour = mean(order_hour_of_day))|>
  pivot_wider(
    names_from = dow,
    values_from = mean_hour
  ) |>
  knitr::kable(digits = 2)
```


# Problem 2 

## 1. Load Data

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

## 2. Data Cleaning 

```{r}
brfss_df = brfss_smart2010 |>
  janitor::clean_names()|>
  rename(state = locationabbr, state_county = locationdesc) |>
  filter(topic == "Overall Health") |>
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) |>
  arrange(response)
```

## 3. Answer Questions

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_df |>
  filter(year == 2002 | year == 2010) |>
  group_by(year, state)|>
  summarize(location = n_distinct(state_county)) |>
  filter(location>=7) |>
  arrange(year, location) |>
  knitr::kable()
```

* The table shows the states that were observed at 7 or more locations in 2002 and 2010. In 2002,CT,FL,NC,MA NJ, and PA were observed at 7 or more location. In 2010, CO, PA, SC, OH, MA, NY, NE, WA, CA, MD, NC, TX, NJ and FL were observed at 7 or more location. 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_df |>
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarize(mean_data_value = mean(data_value, na.rm = TRUE)) |>
  ggplot(aes(x = year, y = mean_data_value, color = state )) +
  geom_line() +
   labs(
    title = "Average Data Value Overtime Across States",
    x = "Year",
    y = "Average Data Value",
    color = "State"
    ) 
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_df |>
  filter(year == 2006 | year == 2010, state == "NY") |>
  ggplot(aes(x = response, y = data_value, color = state_county)) +
  geom_point()+
  facet_grid(. ~ year)+
  labs(
    title = "Distribution of Data Value for Responses Among Locations in NY State",
    x = "Response",
    y = "Data Value",
    color = "Location"
    ) 
```

# Problem 3

## 1. Load, tidy, merge, and otherwise organize the data sets.
  Include all originally observed variables
  Exclude participants less than 21 years of age, and those with missing demographic data; 
  And encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}
nhanes_demo_df = read_csv("./data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names()|>
  mutate(
    sex = ifelse(sex == 1, "male", "female"),
    education = recode(education,
         "1" = "Less than high school", 
         "2" = "High school equivalent", 
         "3" = "More than high school"
         )
    )|>
  mutate(education = factor(education, levels = c("Less than high School", "High school equivalent", "More than high school")))


nhanes_accel_df = read.csv("./data/nhanes_accel.csv")|>
  janitor::clean_names()

nhanes_df = 
  full_join(nhanes_demo_df, nhanes_accel_df) |>
  drop_na(sex | age | bmi | education)|>
  filter(age >= 21)
```




