---
title: "p8105_hw3_tw2918"
output: github_document
date: "2023-10-07"
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1

### 1. Load Data

```{r}
library(p8105.datasets)
data("instacart")
```

### 2.  write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations.

* The dataset contains cleaned and limited data of the “The Instacart Online Grocery Shopping Dataset 2017”, which records theover 3 million online grocery orders from more than 200,000 Instacart users. 
*The dataset contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each `row`/observation in the dataset is a product from an order.
* Some variables reflect key identifications of the orders and customers including `order_id` and `order_id`. There is a single order per consumers in this dataset.
* `eval_set`reflects evaluation set this order belongs in, which is exclusively "`train`" in this dataset. 
* `order_number` reflects the order sequence number for this user. Its value ranges from `r min(instacart$order_number)` to`r max(instacart$order_number)`. The average order_number is `r mean(instacart$order_number)`. 
* Some variables reflect key information of the products, including `product_id` and `product_name`, which aisle and department the prodcuts are purchased from, including  `aisle_id`,  `aisle`,  `department_id` and  `department`, and the order in which each product was added to cart `add_to_cart_order` , from `r min(instacart$add_to_cart_order)` to `r max(instacart$add_to_cart_order)`. 
* `reordered` variable is a binary indicator off this prodcut has been ordered by this user in the past, 0 otherwise. 
* The rest of the varialbes reflect date and time of the order, including the day of the week on which the order was placed `order_dow` ,  the hour of the day on which the order was placed, `order_hour_of_day`, and days since the last order, capped at `30`, `NA` if `order_number=1`. The average days since the last order is `r mean(instacart$order_hour_of_day)`. 

### 3. Answer questions and comments on result

Calculate number of aisle in dataset and which aisles is the most items ordered from

```{r}
mode = function(x, na.rm = FALSE) {
  if(na.rm){ #if na.rm is TRUE, remove NA values from input x
    x = x[!is.na(x)]
  }
  val <- unique(x)
  return(val[which.max(tabulate(match(x, val)))])
}

mode(instacart$aisle_id)
mode(instacart$aisle)
```

* There are `r max(instacart$aisle_id)`  aisles in the dataset, and aisle `r mode(instacart$aisle_id)`, `r mode(instacart$aisle)`aisles is the most items ordered from. 





